---
title: "conditions impacting covid"
author: "Kotomi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

NOTE: I avoided pushing the .csv file that I used in this analysis b/c GitHub says it's too large. Will try and make the data accessible elsewhere but for now they can be downloaded from the link in the project doc

```{r}
library(dplyr)
library(tidyverse)
library(janitor)
```

```{r}
getwd()
```


```{r}
conditions <- read.csv("data/conditions_covid.csv") %>% clean_names()
```

```{r}
summary(conditions)
```
Notes on the raw data:
* the dataset already contains totals grouped by year and month, as indicated by the value in the "group" columns. 
* some potential variables for analyses supported by this data:
    + age group
    + condition/condition group
    + state
    + time period

Overall, across all 3 years, and all age groups, which condition group contributed to the most COVID-19 deaths? Which condition in that group? 
```{r}
conditions %>% 
  filter(group=="By Total" & covid_19_deaths!="NA" & state=="United States") %>% 
  group_by(condition_group) %>% 
  summarise(total_deaths=sum(covid_19_deaths)) %>% 
  arrange(desc(total_deaths))
```
does the listed condition being COVID-19 just mean that COVID itself was the only condition leading to death?

```{r}
conditions %>% 
  filter(condition_group=="Respiratory diseases" & group=="By Total" & covid_19_deaths!="NA" & state=="United States") %>% 
  group_by(condition) %>% 
  summarise(total_deaths=sum(covid_19_deaths)) %>% 
  arrange(desc(total_deaths))
```

Across all 3 years: How do the conditions contributing to deaths from COVID vary by state?

```{r}
conditions %>% 
  filter(group=="By Total" & covid_19_deaths!="NA" & state!="United States") %>% 
  group_by(state, condition_group) %>% 
  summarise(total_deaths=sum(covid_19_deaths)) %>% 
  arrange(desc(total_deaths))
```

Which condition group is most frequently associated with death in age group of 0-24?
```{r}
conditions %>% 
  filter(group=="By Total" & covid_19_deaths!="NA" & state=="United States", age_group=="0-24") %>% 
  group_by(condition_group) %>% 
  summarise(total_deaths=sum(covid_19_deaths)) %>% 
  arrange(desc(total_deaths))
```

Potential idea: if we can standardize date format across case, death, and vaccination data, maybe we can join those datasets and do analysis across them. 




