---
title: "conditions impacting covid"
author: "Kotomi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

NOTE: I avoided pushing the raw data .csv file that I used in this analysis b/c GitHub says it's too large. Will try and make the data accessible elsewhere but for now they can be downloaded from the link in the project doc

```{r}
library(dplyr)
library(tidyverse)
library(janitor)
```

```{r}
getwd()
```


```{r}
conditions <- read.csv("data/conditions_covid.csv") %>% clean_names()
```

```{r}
#write.csv(conditions, file = "conditions_clean.csv", row.names = F)
```


```{r}
summary(conditions)
```
Notes on the raw data:  
* the dataset already contains totals grouped by year and month, as indicated by the value in the "group" columns.  
* some potential variables for analyses supported by this data:  
    + age group  
    + condition/condition group  
    + state  
    + time period  

Overall, across all 3 years, and all age groups, which condition group contributed to the most COVID-19 deaths? Which condition in that group? 
```{r}
conditions %>% 
  filter(group=="By Total" & covid_19_deaths!="NA" & state=="United States") %>% 
  group_by(condition_group) %>% 
  summarise(total_deaths=sum(covid_19_deaths)) %>% 
  arrange(desc(total_deaths))
```
does the listed condition being COVID-19 just mean that COVID itself was the only condition leading to death?

```{r}
conditions %>% 
  filter(condition_group=="Respiratory diseases" & group=="By Total" & covid_19_deaths!="NA" & state=="United States") %>% 
  group_by(condition) %>% 
  summarise(total_deaths=sum(covid_19_deaths)) %>% 
  arrange(desc(total_deaths))
```

Across all 3 years: How do the conditions contributing to deaths from COVID vary by state?

```{r}
conditions %>% 
  filter(group=="By Total" & covid_19_deaths!="NA" & state!="United States") %>% 
  group_by(state) %>% 
  summarise(total_deaths=sum(covid_19_deaths)) %>% 
  arrange(desc(total_deaths))
```

Which condition group is most frequently associated with death in age group of 0-24?
```{r}
conditions %>% 
  filter(group=="By Total" & covid_19_deaths!="NA" & state=="United States", age_group=="0-24") %>% 
  group_by(condition_group) %>% 
  summarise(total_deaths=sum(covid_19_deaths)) %>% 
  arrange(desc(total_deaths))
```

Potential idea: if we can standardize date and state format across case, death, and vaccination data, maybe we can join those datasets and do analysis across them. 

```{r}
unique(conditions$condition)
```


which state had the most COVID-19 deaths involving heart failure?
```{r}
conditions %>% 
  filter(group=="By Total" & covid_19_deaths!="NA" & condition=="Heart failure" & state!="United States") %>% 
  group_by(state) %>% 
  summarise(total_deaths=sum(covid_19_deaths)) %>% 
  ggplot(aes(x=state, y=total_deaths))+
  geom_col()+
  coord_flip()

```

which age group has the most deaths? (all 3 years)
```{r}
conditions %>% 
  filter(group=="By Total" & covid_19_deaths!="NA" & state=="United States") %>% 
  group_by(age_group) %>% 
  summarise(total_deaths=sum(covid_19_deaths)) %>% 
  arrange(desc(total_deaths))
```

```{r}
unique(conditions$state)
```


creating a df with only our states of interest (CA, TX, FL, NY)
```{r}
conditions_states_top4 <- conditions %>% 
  filter(state=="California" | state=="Texas" | state=="Florida" | state=="New York")
  
```

```{r}
#write.csv(conditions_states_top4, file = "conditions_states_top4.csv", row.names = F)
```

