---
title: "conditions analysis v2"
author: "Kotomi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
libraries
```{r}
library("tidyverse")
library("janitor")
library("lubridate")
#install.packages("shiny")
library(shiny)
```
check wd
```{r}
getwd()
```
read csvs
```{r}
totals <- read.csv("data/totals_states_conditions.csv")
monthly <- read.csv("data/monthly_states_conditions.csv")
```
summary functions
```{r}
glimpse(monthly)
```

need to convert dates to date every R session b/c write.csv does not retain date format
```{r}
monthly <- monthly %>% 
  mutate(start_date = as.Date(start_date, format = "%Y-%m-%d"),
         end_date = as.Date(end_date, format = "%Y-%m-%d"))

totals <- totals %>% 
  mutate(start_date = as.Date(start_date, format = "%Y-%m-%d"),
         end_date = as.Date(end_date, format = "%Y-%m-%d"))
```

show trends across all 3 years, all deaths **hiya already doing this, just checking overall trends
```{r}

png(filename = "alldeaths.png", width = 1200, height = 720, units = "px") #exporting this plot as a png, defining file name and dimensions

monthly %>%
  group_by(state, start_date) %>% 
  summarise(total_covid_deaths=sum(covid_19_deaths, na.rm = T),
            .groups = "keep") %>% 
ggplot(aes(x = start_date, y = total_covid_deaths, color = state, group = state)) +
  geom_line(linewidth = 1.2)+
  labs(title = "COVID-19 Deaths in States of Interest", 
       x = "Date", 
       y = "Deaths", 
       color = "State")+
  scale_x_date(labels = scales::date_format("%b %Y"), breaks = "2 months")+
 
  #the following code chunk can be copy-pasted for consistent line-plot formatting
   theme_minimal()+
  scale_colour_brewer(palette = "Paired")+ #indicate the color palette
  theme( #this whole code chunk can be copy-pasted for consistent formatting of title, axis labels, legends
    title = element_text(size=28),
    axis.text.x = element_text(size = 16, angle = 70, hjust = 1),
    axis.text.y = element_text(size = 16),
    legend.position = "bottom",
    legend.text = element_text(size = 20))
```


show trends across all 3 years, associated with circulatory disease
```{r}

png(filename = "allcirculatory.png", width = 1200, height = 720, units = "px")

monthly %>% 
  filter(condition_group=="Circulatory diseases") %>% 
  group_by(state, start_date) %>% 
  summarise(total_covid_deaths=sum(covid_19_deaths, na.rm = T),
            .groups = "drop") %>% 
ggplot(aes(x = start_date, y = total_covid_deaths, color = state, group = state)) +
  geom_line(linewidth = 1.2)+
  labs(title = "COVID-19 Deaths Associated with Circulatory Disease", 
       x = "Date", 
       y = "Deaths", 
       color = "State")+
  scale_x_date(labels = scales::date_format("%b %Y"), breaks = "2 months")+
  theme_minimal()+
  scale_colour_brewer(palette = "Paired")+
  theme( #this whole code chunk can be copy-pasted for consistent formatting of title, axis labels, legends
    title = element_text(size=28),
    axis.text.x = element_text(size = 16, angle = 70, hjust = 1),
    axis.text.y = element_text(size = 16),
    legend.position = "bottom",
    legend.text = element_text(size = 20))
```

trends across all 3 years, associated with obesity
```{r}

png(filename = "allobesity.png", width = 1200, height = 720, units = "px")

monthly %>% 
  filter(condition_group=="Obesity") %>% 
  group_by(state, start_date) %>% 
  summarise(total_covid_deaths=sum(covid_19_deaths, na.rm = T),
            .groups = "drop") %>% 
  
ggplot(aes(x = start_date, y = total_covid_deaths, color = state, group = state)) +
  geom_line(linewidth = 1.2)+
  labs(title = "COVID-19 Deaths Associated with Obesity", 
       x = "Date", 
       y = "Deaths", 
       color = "State") +
  scale_x_date(labels = scales::date_format("%b %Y"), breaks = "2 months") +
  theme_minimal()+
  scale_colour_brewer(palette = "Paired")+
  theme( #this whole code chunk can be copy-pasted for consistent formatting of title, axis labels, legends
    title = element_text(size=28),
    axis.text.x = element_text(size = 16, angle = 70, hjust = 1),
    axis.text.y = element_text(size = 16),
    legend.position = "bottom",
    legend.text = element_text(size = 20))
```

barplot showing percentage of deaths due to particular conditions, by state
```{r}

png(filename = "pctbystate.png", width = 1200, height = 720, units = "px")

totals %>% 
  group_by(state, condition_group) %>% 
  summarise(deaths=sum(covid_19_deaths, na.rm = T),
            .groups = 'drop') %>% 
  ggplot(aes(x=state, y=deaths, fill=condition_group))+
  geom_col(position = position_fill())+
  scale_y_continuous(labels = scales::percent)+
  labs(title = "% of COVID-19 Deaths Associated with Various Conditions, by State",
       x= "State",
       y=NULL,
       fill="Condition Type")+
  
  #the following code chunk can be copy-pasted for consistent barplot formatting
  theme_minimal()+
  scale_fill_brewer(palette = "Paired", labels = function(x) str_wrap(x, width = 40))+ #applies text wrapping in the legend at ~40 characters?
  theme( 
    title = element_text(size=28),
    axis.text.x = element_text(size = 20), 
    axis.text.y = element_text(size = 20),
    legend.position = "right", #looks better on bar plot
    legend.text = element_text(size = 20))

```

barplot showing percentage of deaths due to particular conditions, by age group
```{r}

png(filename = "pctbyage.png", width = 1200, height = 720, units = "px")

totals %>% 
  filter(age_group!="All Ages" &
           age_group!="Not stated") %>% 
  group_by(age_group, condition_group) %>% 
  summarise(deaths=sum(covid_19_deaths, na.rm = T),
            .groups = 'drop') %>% 
  ggplot(aes(x=age_group, y=deaths, fill=condition_group))+
  geom_col(position = position_fill())+
  scale_y_continuous(labels = scales::percent)+
  labs(title = "% of COVID-19 Deaths Associated with Various Conditions, by Age Group",
       x= "Age Group",
       y=NULL,
       fill="Condition Type")+
  theme_minimal()+
  scale_fill_brewer(palette = "Paired", labels = function(x) str_wrap(x, width = 40))+ #applies text wrapping in the legend at ~40 characters?
  theme( 
    title = element_text(size=28),
    axis.text.x = element_text(size = 20),
    axis.text.y = element_text(size = 20),
    legend.position = "right", #looks better on bar plot
    legend.text = element_text(size = 20))
```

barplot showing percentage of deaths due to particular conditions, by year
```{r}

png(filename = "pctbyyear.png", width = 1200, height = 720, units = "px")

monthly %>%  
  group_by(year, condition_group) %>% 
  summarise(deaths=sum(covid_19_deaths, na.rm = T),
            .groups = 'drop') %>% 
  ggplot(aes(x=year, y=deaths, fill=condition_group))+
  geom_col(position = position_fill())+
  scale_y_continuous(labels = scales::percent)+
  labs(title = "% of COVID-19 Deaths Associated with Various Conditions, by Year",
       x= "Year",
       y=NULL,
       fill="Condition Type")+
  theme_minimal()+
  scale_fill_brewer(palette = "Paired", labels = function(x) str_wrap(x, width = 40))+ #applies text in the legend wrapping at ~40 characters?
  theme( 
    title = element_text(size=28),
    axis.text.x = element_text(size = 20),
    axis.text.y = element_text(size = 20),
    legend.position = "right", #looks better on bar plot
    legend.text = element_text(size = 20))
```

breakdown of respiratory diseases associated with covid deaths, by age group
```{r}
totals %>% 
  filter(condition_group=="Respiratory diseases" & 
         age_group!="All Ages" &
           age_group!="Not stated") %>% 
  group_by(age_group, condition) %>% 
  summarise(deaths=sum(covid_19_deaths, na.rm = T),
            .groups = 'drop') %>% 
  ggplot(aes(x=age_group, y=deaths, fill=condition))+
  geom_col(position = position_fill())+
  scale_y_continuous(labels = scales::percent)+
  scale_fill_brewer(palette = "Paired", labels = function(x) str_wrap(x, width = 40))+ #applies text in the legend wrapping at ~40 characters?
  theme_minimal()+
  labs(title = "% Breakdown of Respiratory disease-associated COVID deaths, by Age Group",
       x= "Age Group",
       y=NULL,
       fill="Condition")
```

shiny app time! the idea is to have the user be able to select a condition group another variable (state, age group). shiny app is in a separate R script.

```{r}
unique(totals$condition_group)
```
```{r}
unique(totals$condition)
```


```{r}
names(totals)
```




