---
title: "Weekly United States COVID-19 Cases and Deaths by State"
output: html_document
---

```{r}
library("tidyverse") 
library("dplyr")
library("janitor")

```
```{r}
Weekly <- read_csv("/Users/stephanie/Desktop/BIS15W2025_group13/Stephanie/WeeklyCases.csv") %>% clean_names()
```
```{r}
summary(Weekly)
```
graps of new cases and total deaths. top 5 highest states/lowest states. time variable. see how it increases over weeks and seasons

```{r}
Weekly %>% 
  group_by(state) %>% 
  select(state, new_cases, start_date) %>% 
  filter(state=="CA") %>% 
  mutate(week_number=row_number()) %>% 
  ggplot(aes(x=week_number, y=new_cases))+
  geom_point()+
  labs(title="Weekly New Cases in California 2020-2023",
       x="Weeks",
       y="New Cases")
```

Observations:
- Where are the peaks?
- Do some research, why the sudden peaks?
- What is the highest peak?
- Which seasons/months were the highest peaks? Maybe start of new school session or holidays like christmas, people likely to spend more time outside in close contact. 

Connections with death?
- The weeks with most deaths?
- How much did it collide with new cases

Test to see if save pushes

```{r}
Weekly %>% 
  select(state, tot_deaths) %>% 
  arrange(-tot_deaths) %>% 
  filter!="CA"
```

